---
title: "Homework 3: Bike Sharing Data Wrangling"
author: "Gloria Grace"
output: prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
# Import some functions from lubridate using less confusing names
get_year <- lubridate::year; get_month <- lubridate::month; get_day <- lubridate::day; get_weekday <- lubridate::wday

# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
```


## Load Data

```{r load-data}
# Read the file, specifying data types for each column.
rides_2011 <- read_csv("data/rides_2011.csv.xz", col_types = cols(
  start_time = col_datetime(),
  duration = col_double(),
  rider_type = col_factor()
))
```

## Exercise 1

There are `r nrow(rides_2011)` observation and `r ncol(rides_2011)` variables in this data set.
For each ride in the data frame, it is also given the start time, duration of the ride and its rider type.

## Exercise 2

```{r adding duration-hour}
rides_2011 %>% 
  mutate(duration_hour = duration / 60) %>% 
  arrange(round(duration_hour, digits = 5))
```

By looking at the data, I would assume that the unit that duration uses is minutes.

## Exercise 3

```{r Sum rider types}
rides_2011 %>% 
  group_by(rider_type) %>% 
  summarise(rides = n())
```

## Exercise 4

```{r daily date rides}
daily_rides <- rides_2011 %>% 
  mutate(year = get_year(start_time), 
         month = get_month(start_time), 
         day = get_day(start_time)) %>% 
  group_by(year, month, day) %>% 
  summarise(rides = n())
```


## Exercise 5

In the daily-rides data frame there is `r nrow(daily_rides)` rows or observation. I think it should have 365 because the data is based on the year of 2011 where one year equals 365 days. 

## Exercise 6

```{r daily rides based on rider type}
daily_rides_by_type <- rides_2011 %>%
  mutate(year = get_year(start_time), 
         month = get_month(start_time), 
         day = get_day(start_time)) %>% 
  group_by(year, month, day, rider_type) %>% 
  summarise(rides = n())
```

## Exercise 7

```{r labeling weekdays}
rides_2011 <- rides_2011 %>% 
  mutate(weekday = get_weekday(start_time, label = TRUE)) %>% 
  filter(rider_type != "Unknown") 

```
```{r distribution rides by week}
rides_by_week <- rides_2011 %>% 
  group_by(weekday, rider_type) %>% 
  summarise(rides = median(n()))
  
```

```{r Rides by week plot}
ggplot(data = rides_by_week, mapping = aes(x = weekday, y = rides))+
  geom_boxplot(aes(color = rider_type)) +
  labs(x = "Day of Week",
       y = "Rides",
       color = "Type of Rider", 
       title = "Distributions of rides by day of week") 
```


## Exercise 8

```{r Rides by day of week}
rides_2011 %>% 
  group_by(weekday, rider_type) %>% 
  summarise(median_rides = median(n())) %>% 
  pivot_wider(names_from = rider_type, values_from = median_rides) %>% 
  knitr::kable()
```
