---
title: "Homework 1: Warmup"
author: "Gloria Grace"
output:
  prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r packages, message=FALSE}
library(tidyverse)
library(ggplot2)
```

# Read Data

```{r read-data, message=FALSE}
daily_rides <- read_csv("data/day_simple.csv")
```

# Exercise 1

The data set contains four rows and six rows in total for including the cols().
Inside the cols(), in the four rows, each line has a variable that is assigned to a function. 
The first row represents date, the second row represents the day of the week, the third row represents whether the day is considered a holiday or weekday, or weekend, and the last row represents the number of users that register for a Capital Bikeshare account that day.

# Exercise 2

The following plot shows the total number of rides each day.

```{r rides-by-day}

ggplot(data = daily_rides, mapping = aes(x = date, y = total_rides)) +
  geom_point() +
  geom_smooth() +
  labs(x = "date", y = "number of rides") 
 
```

# Exercise 3

The following plot shows the total number of rides for weekdays vs weekends.

``````{r rides-by-weekday or weekend}

ggplot(data = daily_rides, mapping = aes(x = date, y = total_rides)) +
  geom_point() +
  geom_smooth() +
  labs(x = "date", y = "number of rides") +
  aes(color = workingday)
 
```



# Exercise 4

The number of rides on workdays are much more large than in the weekend, assuming that people use Capital Bikeshares to go to work. 

