---
title: "Linear Model Practice"
author: "STAT 341, Feb. 13, 2023"
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    code_download: true
  pdf_document: default
---

```{r setup, include=FALSE, message=FALSE}
library(faraway)
library(tidyverse)
library(rethinking)
library(CalvinBayes)
knitr::opts_chunk$set(echo = TRUE)
```

### Cake! (Chocolate, of course.) 
An industrious grad student studied the effect of different recipes and baking conditions on the breaking angle of cake (this is the angle to which a slice can be bent before it breaks in two). The data are in package `faraway`.

```{r }
library(faraway)
head(choccake)

choccake
```

## Response and Predictor Variable

Response: breakang
Main predictor: temp

The reason I chose `temp` as my main predictor is because since the grad student is interested in studying the effect between recipe and breaking, I wanted to see whether the recipe affect it. 


## Causal Diagram

```{r}
library(dagitty)
library(CalvinBayes)
```

```{r}
# create the DAG object
causal_diagram_1 <- dagitty("dag{
  temp -> breakang;
  recipe -> breakang;
  batch -> breakang
}")
# plot it
gg_dag(causal_diagram_1)
```

Based on the causal diagram, we think that it would be best that 

## Description of model 

Recall our model is for `household_annual_income` in thousands of Fijian dollars:

$$\text{breaking angle}_i \sim \text{Normal}(\mu_i, \sigma)$$


$$\mu_i = \beta_0 + \beta_1 * \text{temp}_i$$

$$\beta_0 \sim \text{Normal}(30,20)$$
$$\beta_1 \sim \text{Normal}(0.1, 0.05)$$

$$\sigma \sim \text{Normal}(\text{mean}_2 = 10, \text{sd}_2 = 20)$$

## Check at the Data

```{r}
gf_point(breakang ~ temp, data = choccake)
```

## Prior Predictive Check 

```{r}
b0 = rnorm(1,mean = 30, sd = 20)
b1 = rnorm(1,mean = 0.1, sd = 0.05)

simulated_breakang <- b0 + b1 * choccake$temp

gf_point(simulated_breakang ~ temp, data = choccake)
```


## Fit your model 

```{r}
model_descrip <- alist(
  # note variable name has to match actual data variable name
  breakang ~ dnorm(mu, sigma),
  mu ~ b0 + b1 * temp,
  b0 ~ dnorm(30,20),
  b1 ~ dnorm(0.1,0.05),
  sigma ~ dnorm(mean = 10, sd = 20)
)

quap_breakang <- quap(flist = model_descrip,
                          data = choccake)
quap_breakang_sample <- extract.samples(quap_breakang, n = 100)

```


## Graph Posterior(s)

```{r}
gf_dens(~b0, data = quap_breakang_sample)
```

```{r}
gf_dens(~b1, data = quap_breakang_sample)
```

```{r}
gf_dens(~sigma, data = quap_breakang_sample)
```

## Lines

```{r}
gf_point(breakang ~ temp, 
         data = choccake,
         alpha = 0.5) |>
  gf_abline( # draws "line"s given "a" (slope) and "b" (intercept)
    slope = ~b1, # draws one line for each row of psamp
    intercept = ~b0,
    data =  quap_breakang_sample,
    alpha = 0.2, # closer to 0 = more transparent
    linewidth = 0.1 # smaller = thinner
  )
```

## Posterior Predictive Distribution

```{r}
quap_breakang_sample <- quap_breakang_sample |>
  # add row numbers to identify each draw from the posterior
  mutate(sim_id = c(1:nrow(quap_breakang_sample))) |>
  rowwise()  |> # one simulation per draw from posteriors
  mutate(
    # mean_condition is the expected response
    # it's a list with one value per row of original data, `bridges`
    avg_break_ang = list(b0 + b1 * choccake$temp),
    # store predictor data too to make graphs easier later
    temp_new = list(choccake$temp)) |>
  mutate(
    # use sigma to simulate how far datapoints are from expected response 
    sim_condition = list(avg_break_ang +
                           rnorm(n = nrow(choccake), mean = 0, sd = sigma))
    ) |>
  unnest(cols = c(temp_new, avg_break_ang, sim_condition)) |>
  ungroup()
```

```{r}
gf_point(sim_condition ~ temp_new | sim_id,
         data  = quap_breakang_sample |> filter(sim_id <= 10),
         alpha = 0.2)
  # dashed lines show ranges observed in real data
  # (not needed if there are not lower/upper bounds on possible values of response)
  
```

```{r}
gf_point(breakang ~ temp, data = choccake)
```

```{r}
gf_density2d_filled(sim_condition ~ temp_new, 
                    data  = quap_breakang_sample) |>
  gf_theme(scale_fill_grey('')) |> # use greyscale color palette
  gf_refine(guides(fill = FALSE)) |> # don't show color legend
  # (optional) add the real data
  gf_point(breakang ~ temp, data =  choccake,
           size = 0.5, alpha = 0.05) # VERY nearly transparent
```

