---
title: "Stat 245 -- HW ??"
author: "A. Student"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 2.2
    fig_width: 4
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
library(MuMIn)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


## Data

```{r}
age_guess <- read_csv("https://sldr.netlify.app/data/age-guess.csv")
head(age_guess)
```

## Plan a Regression Model

The response variable will be age_guess_error, and the predictors will be actual_age, ethnicity, and gender. 

## Graphics

```{r}
gf_point(age_guess_error ~ actual_age, data = age_guess, alpha = .5)
```


## Fit Model

```{r}
lm_age_guess <- lm(age_guess_error ~ actual_age + ethnicity + gender,data=age_guess)
summary(lm_age_guess)
```


## Model Assessment

```{r}
age_guess <- age_guess |>
  mutate(preds = predict(lm_age_guess),
         resids = resid(lm_age_guess))
```

### Normality of Residuals

```{r}
gf_histogram(~ resids, data = age_guess, bins = 10)
```

The residuals appear to be normally distributed. Our linear model passes this condition.

### Lack of Non-linearity

```{r}
gf_point(resids ~ preds, data = age_guess, alpha = .5)
```

There are no observable trends, so our linear model passes this condition.

### Constant Residual Variance

```{r}
gf_point(resids ~ preds, data = age_guess, alpha = 0.5)
```
The points appear evenly spread around y = 0, so the constant residual variance is good. 


### Independence of Residuals

```{r}
s245::gf_acf(~lm_age_guess) |>
  gf_lims(y = c(-0.5, 0.5))
```
None of the spikes go out of the boundaries.

## Model Selection

```{r}
lm_age_guess <- lm_age_guess |>
  update(na.action = 'na.fail')
age_guess_dredge <- dredge(lm_age_guess, rank = 'BIC')
age_guess_dredge
```

## Prediction Plot

### Hypothetical Dataset
```{r}
age_guess_pred_data <- expand.grid(
              actual_age = seq(from = 5, by = 1, to = 80),
                             ethnicity = 'Caucasian/White',
                             gender = 'Female')
```

### Make predictions from fitted model
```{r}
age_guess_preds <- predict(lm_age_guess, newdata = age_guess_pred_data, se.fit = TRUE)
```

### Make predictions
```{r}

age_preds <- predict(lm_age_guess,
newdata = age_guess_pred_data, 
se.fit = TRUE)
```


```{r}
age_guess_pred_data <- age_guess_pred_data %>% 
  mutate(pred = age_preds$fit, pred.se = age_preds$se.fit)
```

```{r}
age_guess_pred_data <- age_guess_pred_data %>% mutate(CI_lower = pred - 1.96*pred.se, CI_upper = pred + 1.96*pred.se)
```


```{r}
gf_line(pred ~ actual_age, data=age_guess_pred_data) %>% 
    gf_ribbon(CI_lower + CI_upper ~ actual_age)
```



Our best model is the one with actual age and gender. It passed all of the model assessment conditions.














  