---
title: "MCMC"
author: "STAT 341, Mar 24, 2023"
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    code_download: true
  pdf_document: default
---

```{r setup, include=FALSE, message=FALSE}
library(faraway)
library(tidyverse)
library(rethinking)
library(CalvinBayes)
library(tidybayes)
library(rstan)
knitr::opts_chunk$set(echo = TRUE)
```

### data
```{r}
library(rethinking)
data("Wines2012")
Wines2012 <- Wines2012 |>
  # make a standardized version of the score variable
  mutate(score_std = as.numeric(scale(score)),
         # categorical versions of some variables originally coded 0/1
         wine.origin = ifelse(wine.amer == 1, 'USA', 'Other'),
         judge.nationality = ifelse(judge.amer == 1, 'USA', 'France'),
         # numeric index versions of the same categorical variables
         origin.ix = as.numeric(factor(wine.origin)),
         judge.ix = as.numeric(factor(judge.nationality))
         )

```
```{r}
stan_data <- compose_data(Wines2012)
```

```{r}
Wines2012 <- Wines2012 |>
  mutate(Origin_Judge = interaction(wine.origin, judge.nationality),
         Origin_Judge_ix = as.numeric(Origin_Judge)) |>
  drop_na(Origin_Judge, score_std)
glimpse(Wines2012)
```


### Description

$$\text{score_std}_i \sim \text{Normal}(\mu, \sigma)$$

For some reason this won't work (rmd won't knit)even though I put dollar signs on them.
## mu_i = beta_1 text{[Origin_Judge_ix]}

$$\mu_i = \beta_1$$
$$\beta_1 \sim \text{Normal}(0.1, 0.25)$$
$$\sigma \sim \text{Normal}(\text{mean}_2 = 0, \text{sd}_2 = 1)$$

### Defining the interaction

The interactions of the variable `wine.origin` and `judge.nationality` means that the effect of `wine.origin` variable on the response variable is different at different values of the `judge.nationality` variable. Adding a term to the model in which the two predictor variables are multiplied tests this.

### Priors
Rationale:

My rational is that I pick a number after looking at the levels of the variable which is there is 4 options. 

### Fit

```{r}
levels(Wines2012$Origin_Judge)
```

```{r}
wine_model_list <- alist(
  score_std ~dnorm(mu, sigma),
  mu <- b1[Origin_Judge_ix],
  b1[Origin_Judge_ix] ~ dnorm(0, 0.25),
  sigma ~ dlnorm(0,1)
)

```
```{r}
wine_int_model <- quap(
  flist = wine_model_list,
  data = Wines2012
)
```

```{r}
wine_ps <- extract.samples(wine_int_model) |>
  data.frame()
glimpse(wine_ps)
```

```{r}
wine_ps_plot <- pivot_longer(wine_ps, cols = everything())
glimpse(wine_ps_plot)
```


```{r}
gf_dens(~value, data = wine_ps_plot) |>
  gf_facet_wrap(~name, scales = "free_y")
```

```{r}
gf_dens(~value, color = ~name, data = wine_ps_plot, size = 2)
```
```{r}
levels(Wines2012$Origin_Judge)
```

### Interpret

Based on the plot, when the origin is USA and the judge is from france, it has the lowest value for the score_std whereas the highest value goes to the origin other than USA and USA judge. 

## Stan Program

```{r}
stan_program <- '
data {
// number of observations
int<lower=1> n;
// response
vector[n] score_std;
// categorical predictor,"Other.France" or "USA.France" or "Other.USA" or   "USA.USA" 
array[n] int Origin_Judge;

}
parameters {
// std of response, single continuous value
real<lower=0> sigma;
// vector of two numeric values: intercepts for Other.France, USA.France, Other.USA, and USA.USA 
vector[4] a;
// vector of two numeric values: slopes for Other.France, USA.France, Other.USA, and USA.USA
vector[4] b;
}

model {
// vector of n values: expected score_std for each observation
vector[n] mu;
// loop over the n cases in the dataset to estimate mu_i values
for (i in 1:n) {
// expected score_std is region-specific intercept +
mu[i] = a[Origin_Judge[i]];
}
// prior for both intercepts
a ~ normal(0, 0.25);
// prior for sigma
sigma ~ normal(0,1);
// defining likelihood in terms of mus and sigma
score_std ~ normal(mu, sigma);
}
'
```


DOESN'T WORK 

```{r}
# fit the model defined above to the stan_data on ruggedness and GDP
wine_model <- stan(model_code = stan_program, data = stan_data)


```

```{r}
wine_model <- stan(model_code = stan_program,
                     data = stan_data,
                     iter = 2000,
                     warmup = floor(iter/2),
                     chains = 4,
                     control = list(adapt_delta = 0.8))

```

