---
title: "Stat 245 -- Search Count Model"
author: "Gloria Grace"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
classoption: landscape
header-includes:
- \usepackage[fontsize=20pt]{scrextend}
---


```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
library(glmmTMB)
library(DHARMa)

knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      fig.width = 7, 
                      fig.height = 4)

theme_set(theme_minimal(base_size = 16))
```


## Data 

```{r}
search <- read_csv('https://sldr.netlify.app/data/election_searches.csv')
```

## Plan

For this model, the predictors that I want to include which I think is relevant to the number of internet searches someone does per week (searches) are Age, Education, and Sex. One of the reasons I included both age and education is because I expect that the higher education, the higher the number of searches it will be. For Age, I expect people in the 30s below might have higher search count compares to people in their 60s and above. 

## Fit

```{r}
search_nb1 <- glmmTMB(Searches ~ Age + Education + Sex, data = search,
                      family = nbinom1(link = 'log'))
```

```{r}
search_nb2 <- glmmTMB(Searches ~ Age + Education + Sex, data = search,
                      family = nbinom2(link = 'log'))
```

Check which is better: 
```{r}
AIC(search_nb1, search_nb2)
```


## Model Assessment

Scaled Residuals

```{r}
nb2_sim <- simulateResiduals(search_nb2)
plotResiduals(nb2_sim,
              quantreg = FALSE)
```

```{r}
search <- search |>
  mutate(nb2_pred = rank(predict(search_nb2,
                                 type = 'response')),
         nb2_scaled_resid = nb2_sim$scaledResiduals)
```

```{r}
gf_point(nb2_scaled_resid ~ nb2_pred,
         data = search) |>
  gf_labs(y = 'Scaled Resid.', x = 'Predicted Search')
```




  