---
title: "Stat 341 -- Homework 2"
author: "Gloria Grace"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable() for tables
library(patchwork)   # for combining plots

# data wrangling
library(tidyverse)

# several packages for bayesian stuff -- more may come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

## SR 3E1

```{r}
mean(samples < 0.2)
```

## SR 3E2

```{r}
mean(samples > 0.8)
```

## SR 3E3

```{r}
mean(between(samples, 0.2, 0.8))
```

## SR 3E4

```{r}
quantile(samples, probs = 0.2)
```

## SR 3E5

```{r}
quantile(samples, probs = 0.8)
```

## SR 3E6

```{r}
HPDI( samples , prob=0.66)
```

## SR 3E7

```{r}
PI( samples , prob=0.66)
```

## SR 3M1

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 ) 
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
```

```{r}
tibble(p = p_grid, posterior = posterior) %>%
  ggplot(aes(x = p, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(x = "Proportion of Water", y = "Posterior Density")
```

## SR 3M2

```{r}
set.seed(101)
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
HPDI(samples, prob = 0.9)
```

## N1

```{r read data}
movielens <- read_csv('https://sldr.netlify.app/data/movielens.csv')
```

## Part A

The quantity that I am working on to estimate is the proportion of the movie that is comedy. 


```{r}
table(movielens$comedy)
```

```{r}
total_comedy <- 773
total_Ncomedy <- 1172
```


## Part B 
Choose a prior that adequately expresses your prior knowledge about the situation, before peeking at the data. Explain your choice. Make a sketch (by hand) or a graph (in R) of the prior, with appropriate axis titles (y axis need not be perfectly to scale)

I don't fully understand what I should input as my shape 1 and 2 for my beta distribution. 

```{r}
my_beta_prior <- tibble(p_grid = seq(from = 0, to = 1, length.out = 1000),
                        prior_density = dbeta(p_grid, shape1 = 10, shape2 = 4))
gf_line(prior_density ~ p_grid, data = my_beta_prior)
```


## Part C 

fixed:
```{r}
grid_movie2 <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = 100000),      # define grid
         dbeta(p_grid, shape1 = 10, shape2 = 4)) |>                                       # define prior
  mutate(likelihood = dbinom(total_comedy, 
                             size = total_comedy  + total_Ncomedy, 
                             prob = p_grid)) |>  # compute likelihood at each value in grid
  mutate(unstd_posterior = likelihood * prior) |>             # compute product of likelihood and prior
  mutate(posterior = unstd_posterior / sum(unstd_posterior))   # standardize the posterior, so it sums to 1


# to peek at the results table
glimpse(grid_movie2)
```


```{r}
grid_movie <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = 100000),      # define grid
         prior  = 1) |>                                       # define prior
  mutate(likelihood = dbinom(total_comedy, 
                             size = total_comedy  + total_Ncomedy, 
                             prob = p_grid)) |>  # compute likelihood at each value in grid
  mutate(unstd_posterior = likelihood * prior) |>             # compute product of likelihood and prior
  mutate(posterior = unstd_posterior / sum(unstd_posterior))   # standardize the posterior, so it sums to 1


# to peek at the results table
glimpse(grid_movie)
```


## Part D 
Interpret and present your posterior in at least 2 ways. One of them should be a graph (not a number or interval). At least one of them should require sampling from the posterior as a preparatory step. Include a paragraph explaining your choice (why did you choose these 2 particular ways to show/present the posterior)?

```{r}
grid_plot <- gf_point(posterior ~ p_grid, 
                           data = grid_movie) |>
  # this is optional -- adds a line in addition to  the dots
  gf_line(color = 'grey44', alpha = 0.5) |>
  gf_labs(x = 'Possible values of Comedy',
          y = 'Posterior Probability')

grid_plot
```

```{r}
grid_plot |> 
  gf_lims(x = c(0.35, 0.45))
```

Fixed:

```{r}
grid_plot2 <- gf_point(posterior ~ p_grid, 
                           data = grid_movie2) |>
  # this is optional -- adds a line in addition to  the dots
  gf_line(color = 'grey44', alpha = 0.5) |>
  gf_labs(x = 'Possible values of Comedy',
          y = 'Posterior Probability')

grid_plot2
```

```{r}
grid_plot2 |> 
  gf_lims(x = c(0.35, 0.45))
```
## Part E 

Based on my work, I learned that the probability of the movies given that it is a comedy genre is 3e-04 where the possible value of comedy is below 0.4.

I am still confused about what does it mean, since it is binary, I understand there is probability of comedy and not comedy but why is the probability 3e-04? what does the possible values of comedy represent or what is P grid? I think I know what prior and posterior based on what I learn in class but I feel like classes exercise where we have to do our own does not help my understanding that much since a lot of people ask question to. 

## Part F 

I do wonder how the data is obtained, out of all the movies that were picked, how were they pick. I do think that the data or the question that I answer would appear differently if the data were obtained differently too. 




