---
title: "Diversity & Innovation in Academia"
author: "Gloria Grace"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
classoption: landscape
header-includes:
- \usepackage[fontsize=20pt]{scrextend}
---

```{r setup, include=FALSE}
require(tidyverse)
require(mosaic)
require(ggformula)
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      fig.width = 7, 
                      fig.height = 4)

theme_set(theme_minimal(base_size = 16))
```

## Data

```{r}
pi <- read_csv('https://sldr.netlify.app/data/phd_innovation.csv')
```

## Fit Linear Regression

```{r}
mlr <- lm(uptake ~ prop_same_sex + sex, data = pi )
```

## Model Summary

```{r}
summary(mlr)
```
Equation: 

uptake = 0.81 + 0.95 prop_same_sex + 0.04 Imale + $\epsilon$,

$\epsilon$ $\sim$ N(0, 0.64), Imale is 1 if pi is male, 0 otherwise. 


## Making a Prediction Plot

### new prediction data

```{r}
fake_data <- expand.grid(prop_same_sex = seq(from = 0.2, 
                                             to = 0.9,
                                             by = 0.01),
                         sex = c("Male"))
```

### SE Predictions

```{r}
preds <- predict(mlr, 
                 newdata = fake_data, 
                 se.fit = TRUE)
```

```{r}
fake_data <- fake_data %>% 
  mutate(pred = preds$fit, 
         pred.se = preds$se.fit)
```
### Convert from SE to CI

```{r}
fake_data <- fake_data %>% 
  mutate(CI_lower = pred - 1.96*pred.se, 
         CI_upper = pred + 1.96*pred.se)
```

## Prediction Plot 

```{r}
gf_line(pred ~ prop_same_sex, 
         data = fake_data) %>% 
  gf_labs(y = 'Predicted\n uptake') %>% 
  gf_ribbon(CI_lower + CI_upper ~ prop_same_sex)
```

For my fitted model, I chose uptake as my response because I want to investigate other factors with it and both sex and prop_same_sex as my predictors. I chose my model by looking at the data set and understanding what each variable represents. 

From the prediction plot, I can take a conclusion that through this prediction, a measure of how enthusiastically innovative ideas are taken up or uptake are taken higher as proportion of peers who are the same sex also goes higher. 
  