---
title: "Guided Exercise 12 - Databases"
output: 
  html_document: default
---

---
title: "Guided Exercise 12 - Databases"
output: 
  tufte::tufte_html:
    css: ../ex.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
theme_set(theme_bw())
options(scipen = 5) # encourage metrics to print in fixed-point notation
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

```{r}
library(tidyverse)
library(mdsr)
library(RMySQL)
db <- DBI::dbConnect(RMariaDB::MariaDB(), dbname = "airlines", host = "mdsr.cdc7tgkkqd0n.us-east-1.rds.amazonaws.com", 
    user = "mdsr_public", password = "ImhsmflMDSwR")
```

And we'll get references to two specific tables we'll be working with.

```{r}
flights <- tbl(db, "flights")
carriers <- tbl(db, "carriers")
```

```{r}
flights %>% 
  filter(year == 2016) %>% 
  show_query()
```

```{r}
nrow(flights)
```

## Exercise 1

The SQL expression of filter 2016 year as From the data, where 2016 is the year.

```{r query-finding-rows}
flights %>% 
  filter(year == 2016, month == 11, distance > 1000) %>% 
  summarize(
    n = n(),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE)
  ) %>% show_query()
```

```{r finding-n-rows}
flights %>% 
  filter(year == 2016, month == 11, distance > 2000) %>% 
  summarize(n = n(), mean_dep_delay = mean(dep_delay, na.rm = TRUE))
```

## Exercise 2 

SQL express n as count and mean as average. 


## Exercise 3

The years of data that are stored in the flights table is 2010 - 2017.

```{r check-years-stored-in-data}
flights %>% 
  summarize(min(year, na.rm = TRUE),
            max(year, na.rm = TRUE))
```

## Exercise 4

```{r n-flights-arrived-in-JFK}
flights %>% 
  filter(year == 2016, month == 11, dest == 'JFK') %>% 
  summarize(n = n())
    

```

## Exercise 5

```{r n-flights-by-carrier}
flights %>% 
  filter(year == 2016, month == 11, dest == 'JFK') %>% 
  group_by(carrier) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

```

## Exercise 6

SQL express group by as GROUP BY and arrage as ORDER BY.

```{r SQL-expression-group-by-and-arrange}
flights %>% 
  filter(year == 2016, month == 11, dest == 'JFK') %>% 
  group_by(carrier) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  show_query()

```

## Exercise 7

```{r fraction-flights-on-time}
flights %>% 
  filter(dest == 'JFK') %>% 
  group_by(carrier) %>% 
  summarize(n = n(), frac_ontime = sum(arr_delay <= 15)) %>% 
  arrange(desc(n))
  

```

## Exercise 8

